Formalizing Stack Safety as a Security Property\ifcameraready\else\vspace{-0.75cm}\fi

### 2023-05-25 submission5/core.tex
Last minute nits
Fixed typo

\begin{figure*}
  \begin{subfigure}{0.48\textwidth}
    \judgmentbr[~AllocF]
        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
        {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
        {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}

@@

    \judgmentbr[~Dealloc]
        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \object(V)\)}
        {\(V' = V \llbracket \addr \mapsto \unsealed | \addr \in \components \rrbracket\)}
        {\(Op ~ \mach ~ (\mathbf{dealloc} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}

    \caption{Memory Allocation}
    \label{sfig:publicalloc}
  \end{subfigure}
  \begin{subfigure}{0.52\textwidth}

    \[\mathit{push} ~ V ~ \overline{\reg} ~ K \triangleq \lambda k .
    \begin{cases}
      \unsealed & \textnormal{if } k \in CLR \\
      \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
      \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } \\
      & k \in \object(V) - \components \\
      V(k) & \textnormal{otherwise} \\
    \end{cases}      
    \]

    \[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
    \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]

    \judgmenttwo[~Call]
        {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ \components\)}
        {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
          ~ (V,\sigma) = (V',V::\sigma)\)}

@@

    \judgmenttwo[~Tailcall]
        {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ \components\)}
        {\(Op ~ \mach ~ (\mathbf{tail call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
          ~ (V,\sigma) = (V',\sigma)\)}

@@

For example, when loading from a stack location, {\em Depth Isolation}
needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
that restriction by omitting the check (bug {\em
  LOAD\_NO\_CHECK\_DI}).
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraint causes violations to the integrity
property (bug {\em STORE\_NO\_CHECK}).

In additional intentional mutations, our testing catches errors in our
own implementation of the enforcement mechanism, including one interesting
bug where the initial function's frame included space allocated for its
return address, but this uninitialized (and therefore \(\tagNoDepth\)-tagged)
space was treated as private data but left unprotected. We added this to
our set of mutations as {\em HEADER\_NO\_INIT}.
%(the original was written in C++
%and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
%we re-implemented it in Coq targeting RISC-V).  These errors range
%from trivial typos to ones that require an intriguingly complex setup
%to provoke.  The most interesting bug (included in the table as row
%{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
%sequence %/policy combination\apt{??}
%did not initialize all locations for the
%newly allocated stack frame correctly, but left some of them as
%$\tagNoDepth$. This allowed for a potential integrity violation, but
%only if a rather complicated sequence of events occured.
%The smallest counterexample requires calling a function {\tt f},
%which fails to initialize some of its frame during the blessed sequence,
%but writes into an uninitialized location $l$ later, treating \(l\) as outside
%the stack. Then {\tt f} calls a further function {\tt g} (which should have
%the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
%which is allowed because the enforcement mechanism still has
%$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
%  and the write from {\tt f} was already a violation.}
%}

@@

\subsection{Results}

The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
\cref{tab:bug-table}, along with the average number of tests
it took to find the failure. Experiments were run in a desktop
machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.

For LTC, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
fails in testing, in cases where data is leaked between sequential calls.
To round out our mutation testing we also check {\em LOAD\_NO\_CHECK},
equivalent to its counterpart in depth isolation,
and a version where stores succeed but fails to propagate the PC tag, {\em STORE\_NO\_UPDATE}.
It turns out that {\em PER\_DEPTH\_TAG} is a comparatively subtle bug,
taking longer to catch.

>>>>

\begin{figure*}[h]
  \begin{subfigure}{\textwidth}
    \judgmentbr[~AllocF]
        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap \unsealed(V)\)}
        {\(V' = V \llbracket \addr \mapsto \object \mid \addr \in \components \rrbracket\)}
        {\(Op ~ \mach ~ (\mathbf{alloc} ~ \mathbf{f} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}

@@

    \judgmentbr[~Dealloc]
        {\(\components = \mathit{range} ~ \SP ~ \mathit{off} ~ \mathit{sz} ~ \mach \cap (\object(V) \cup \public(V))\)}
        {\(V' = V \llbracket \addr \mapsto \unsealed | \addr \in \components \rrbracket\)}
        {\(Op ~ \mach ~ (\mathbf{dealloc} ~ (\mathit{off, sz})) ~ (V,\sigma) = (V',\sigma)\)}

    \caption{Memory Allocation}
    \label{sfig:publicalloc}
  \end{subfigure}
  \begin{subfigure}{\textwidth}

    \begin{minipage}{0.4\textwidth}
      \[\mathit{push} ~ V ~ \overline{\reg} ~ K \triangleq \lambda k .
      \begin{cases}
        \unsealed & \textnormal{if } k \in CLR \\
        \public & \textnormal{if } k \in \overline{\reg_{\mathit{args}}} \\
        \sealed & \textnormal{if } k \in \WORDS \textnormal{ and } \\
        & k \in \object(V) - \components \\
        V(k) & \textnormal{otherwise} \\
      \end{cases}      
      \]

      \[\mathit{passed} ~ \overline{sa} ~ \mach \triangleq \bigcup_{(\reg,\mathit{off},\mathit{sz}) \in \overline{sa}}
      \mathit{range} ~ \reg ~ \mathit{off} ~ \mathit{sz} ~ \mach\]
    \end{minipage}
    \begin{minipage}{0.6\textwidth}    
      \judgmenttwo[~Call]
        {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ \components\)}
        {\(Op ~ \mach ~ (\mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
          ~ (V,\sigma) = (V',V::\sigma)\)}

@@

    \judgmenttwo[~Tailcall]
        {\(\components = \mathit{passed} ~ \overline{sa} ~ \mach\)}
        {\(V' = \mathit{push} ~ V ~ \overline{\reg_{args}} ~ \components\)}
        {\(Op ~ \mach ~ (\mathbf{tail call} ~ \addr_{target} ~ \overline{\reg_{args}} ~ \overline{sa})
          ~ (V,\sigma) = (V',\sigma)\)}
    \end{minipage}

@@

For example, when loading from a stack location, {\em Depth Isolation}
needs to enforce that the tag on the location being read
is $\tagStackDepth{n}$ for some number $n$ and that the tag of the
current $\PCname$ is $\tagPCDepth{n}$ for the same depth $n$. We can relax
that restriction by omitting the check (bug {\em
  LOAD\_NO\_CHECK}).
%
Similarly, when storing to a stack location, the correct micro-policy
needs to ensure that the tag on the memory location is either
$\tagNoDepth$ or has again the same depth as the current $\PCname$
tag. Relaxing that constraint causes violations to the integrity
property (bug {\em STORE\_NO\_CHECK}).

In additional intentional mutations, our testing catches errors in our
own implementation of the enforcement mechanism, including one interesting
bug where the initial function's frame included space allocated for its
return address, but this uninitialized (and therefore \(\tagNoDepth\)-tagged)
space was treated as private data but left unprotected. We added this to
our set of mutations as {\em HEADER\_NO\_INIT}.

For LTC, the original micro-policy, implemented as {\em PER\_DEPTH\_TAG},
fails in testing, in cases where data is leaked between sequential calls.
To round out our mutation testing we also check {\em LOAD\_NO\_CHECK},
equivalent to its counterpart in depth isolation,
and a version where stores succeed but fails to propagate the PC tag, {\em STORE\_NO\_UPDATE}.

The mean-time-to-failure (MTTF) and average number of tests for various bugs can be found in
\cref{tab:bug-table}, along with the average number of tests
it took to find the failure. Experiments were run in a desktop
machine equipped with i7-4790K CPU @ 4.0GHz with 32GB RAM.
%(the original was written in C++
%and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
%we re-implemented it in Coq targeting RISC-V).  These errors range
%from trivial typos to ones that require an intriguingly complex setup
%to provoke.  The most interesting bug (included in the table as row
%{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
%sequence %/policy combination\apt{??}
%did not initialize all locations for the
%newly allocated stack frame correctly, but left some of them as
%$\tagNoDepth$. This allowed for a potential integrity violation, but
%only if a rather complicated sequence of events occured.
%The smallest counterexample requires calling a function {\tt f},
%which fails to initialize some of its frame during the blessed sequence,
%but writes into an uninitialized location $l$ later, treating \(l\) as outside
%the stack. Then {\tt f} calls a further function {\tt g} (which should have
%the effect of sealing $l$ for integrity purposes). {\tt g} attempts to write to $l$,
%which is allowed because the enforcement mechanism still has
%$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
%  and the write from {\tt f} was already a violation.}
%}

@@
